Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2022-01-11 manuscript/manuscript.Rmd
update manuscript formatting, add author summary

---
title: "Comparing human and model-based forecasts of COVID-19 in Germany and Poland"
output: 
  bookdown::pdf_document2:
    extra_dependencies: ["float"]
    keep_tex:  true
    toc: false
bibliography:
  - references.bib
  - GermanPolishpaper.bib
csl: apa.csl
---

@@

>>>>

---
title: "Comparing human and model-based forecasts of COVID-19 in Germany and Poland"
header-includes:
    - \usepackage{lineno}
    - \linenumbers
output: 
  bookdown::pdf_document2:
    extra_dependencies: ["float"]
    keep_tex:  true
    toc: false
linestretch: 2
bibliography:
  - references.bib
  - GermanPolishpaper.bib
csl: vancouver.csl
---

@@

# Author summary

Mathematical models of COVID-19 have played a key role in informing governments across the world. While mathematical models are informed by our knowledge of infectious disease dynamics, they are ultimately developed and iteratively adjusted by the researchers and shaped by their subjective opinions. To investigate what modelling is able to add beyond the subjective opinion of the researcher alone, we compared human forecasts with model-based predictions of COVID-19 cases and deaths submitted to the so-called German/Polish Forecast Hub (which collates a variety of models from a range of teams).  
We found that our human forecasts consistently outperformed an aggregate of all available model-based forecasts when predicting cases, but not when predicting deaths. Our findings suggest that human insight may be most valuable when forecasting highly uncertain quantities, which depend on many factors that are hard to model using equations, while mathematical models may be most useful in settings like predicting deaths, where leading indicators with a clear connection to the target variable are available. This potentially has very relevant policy implications, as agencies informing policy-makers could benefit from routinely eliciting human forecasts in addition to model-based predictions to inform policies.
